---
title: "TTF PM Tool Input and Output Specification for PortfolioOptimization Package"
output: pdf_document
params:
  strategyname: "ttf"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Background

FI Tech and FIQuant are collaborating on a tool to facilitate the daily portfolio management process of Total Return Fund (TTF).  This tool will leverage a systematic process for portfolio optimization and modeling developed by the FIQuant based on the R package PortfolioOptimization.  This document provides the input (InputObject) and output (OutputObject) object specifications for the top level function, generateOptimalPortfolios, that the PM Tool will call.  

#### Top level function call using the FI Quant FIPortfolioOptimization Package

The top level function call is as follows:

OutputObject <- generateOptimalPortfolios(InputObject)

This function should be run on the fly as users modify any inputs or constraints through the PM Tool.  


```{r, echo = FALSE,include = FALSE}

#library(RJSONIO)
#library(here)
library(knitr)
library(kableExtra)
#source("createSyntheticTTFMasterData.R")

StrategyName <- "ttf"
OptimizationType <- "strategic"
DataFrequency <- "M"
dtsRatio <- 2
TEV <- 100
Confidence <- 0.5
  
cashSectors <- c('govRel', 'tips', 'corp','mbs','abs','cmbs', 'rmbs', 'usHy','levL','clo','emd','emc','eml')
synthetics <- c('igCdx','hyCdx','mainCdx','xoCdx','emCdx', 'SnrFin')
ratesSectors <- c('krd6M','krd2','krd5','krd10','krd20','krd30')

RiskBuckets <- list()
RiskBuckets$cashSectors <- cashSectors
RiskBuckets$ratesSectors <- ratesSectors
RiskBuckets$synthetics <- synthetics

N <- sum(as.numeric(lapply(RiskBuckets,length)))

benchmarkName <- 'Agg'

CovMatrixOrder <- c(cashSectors, synthetics, ratesSectors,benchmarkName)

CovMatrix <- matrix(rexp((N+1)*10, rate=.1), ncol=N+1, nrow = N+1)

Views <- data.frame(matrix(rnorm(N),N,1))
rownames(Views) <- CovMatrixOrder[-(N+1)]
colnames(Views) <- 'ER'

wCashSectors <- rnorm(mean = 0, sd = 0.2, n=length(cashSectors))
wCashSectorsT <- wCashSectors + rnorm(mean = 0, sd = 0.02, n=length(cashSectors))
wSynthetics <- rnorm(mean = .05, sd = 0.1, n=length(synthetics))
wSyntheticsT <- wSynthetics
wRatesSectors <- rnorm(mean = 0, sd = 0.05, n=length(ratesSectors))
wRatesSectorsT <- wRatesSectors + rnorm(mean = 0, sd = 0.02, n=length(ratesSectors))
wCashSectorsBench <- c(0, 0, .27, .27, .01, .02, 0, 0, 0, 0, 0, 0, 0) #fill this in
wSyntheticsBench <- rep(0, length(synthetics))
wRatesSectorsBench <- c(0, .02, .25, .12, .03, .01) #fill this in

df.currentWgt <- data.frame(c(wCashSectors, wSynthetics, wRatesSectors))
colnames(df.currentWgt) <- 'currentWgt'
df.benchmarkWgt <- data.frame(c(wCashSectorsBench, wSyntheticsBench, wRatesSectorsBench))
colnames(df.benchmarkWgt) <- 'benchmarkWgt'
df.targetWgt <- data.frame(c(wCashSectorsT, wSyntheticsT, wRatesSectorsT))
colnames(df.targetWgt) <- 'targetWgt'

rownames(df.currentWgt) <- c(cashSectors, synthetics, ratesSectors)
rownames(df.benchmarkWgt) <- rownames(df.currentWgt)
rownames(df.targetWgt) <- rownames(df.currentWgt)


#load("TTF/InputObject.RData")
#createSyntheticTTFMasterData()
load("TTF/TTFMasterData.RData") #MasterData
ActualCloseAnalytics <- MasterData
dt.aggDly <- trpscda::generate_aggregates("ssag")
dt.aggMly <- dt.aggDly

#InputObject Aggregation
InputObject <- list()
InputObject$StrategyName <- StrategyName #always ttf
InputObject$RiskBuckets <- RiskBuckets
InputObject$Benchmark <- benchmarkName
InputObject$TEV <- TEV
InputObject$dtsRatio <- dtsRatio
InputObject$ActualCloseAnalytics <- ActualCloseAnalytics
InputObject$CovMatrixOrder <- CovMatrixOrder
InputObject$CovMatrix <- CovMatrix
InputObject$Views <- Views
InputObject$Confidence <- Confidence
InputObject$CurrentWgt <- df.currentWgt
InputObject$TargetWgt <- df.targetWgt
InputObject$BenchmarkWgt <- df.benchmarkWgt
InputObject$HistoricalDaily <- dt.aggDly
InputObject$HistoricalMonthly <- dt.aggMly

save(InputObject, file = "TTF/InputObject.RData")

Notes <- list()
Notes$StrategyName <- "Must be ttf"
Notes$RiskBuckets <- "List with following fields: cashSectors, ratesSectors, synthetics "
Notes$Benchmark <- "Must be Agg"
Notes$TEV <- "Must be greater than 0"
Notes$dtsRatio <- "Must be greater than 0"
Notes$ActualCloseAnalytics <- "Contains all close of day analytics for actual subaccounts (and index where actual not available)"
Notes$CovMatrixOrder <- "Contains all risk buckets + benchmark in order shown in cov matrix"
Notes$CovMatrix <- "Must be NxN, where N is length sum all 3 types of RiskBuckets"
Notes$Views <- "Must be length N.  Lack of view indicated by NA"
Notes$Confidence <- "Must be between 0 and 1"
Notes$CurrentWgt <- "Current net actual weights for all risk buckets.  Must be length N"
Notes$TargetWgt <- "Current net target weights for all risk buckets.  Must be length N"
Notes$BenchmarkWgt <- "Current benchmark weights for all risk buckets.  Must be length N"
Notes$HistoricalDaily <- "Historical returns and analytics daily data"
Notes$HistoricalMonthly <- "Historical returns and analytics monthly data"

fieldNames <- attributes(InputObject)$name
expectedType <- as.character(unlist(lapply(InputObject, class)))
expectedType <- expectedType[-7] #remove data.table because it appears twice for MasterData
notes <- as.character(unlist(Notes))

dt.InputObjectSpec <- cbind(fieldNames,expectedType, notes)
cNames <- c("Field Name", "Field Type", "Additional Notes")

```

## InputObject Specification used by generateOptimalPortfolios

The table below contains the fields required by generateOptimalPortfolios in the InputObject:

```{r InputsTable,echo=FALSE,include=TRUE}

knitr::kable(dt.InputObjectSpec,caption = 'Expected Fields in InputObject',col.names = cNames) %>%
  kableExtra::kable_styling(font_size = 8.5)


```

This InputObject should be stored with each saved instance of the PM Tool use.  In case the process to generate the InputObject fails, the InputObject from the previous day can be loaded instead, with the user notified of the exception.  Then the issue is resolved, the user should be notified that the new InputObject has loaded and the model can be rerun with the most current data.

An example InputObject with illustrative (not actual) data is also provided along with this document.  As a note, the historical data represented is illustrative only.  It is proxied by the trpscda::generate_aggregates("ssag") call, which will differ from what is described below.

```{r blank,echo=FALSE,include=TRUE}

```
\newpage

#### TTF Sectors and Risk Buckets

The following sectors/risk buckets and required data elements are used in different areas of the covarience estimation, portfolio optimization or output visualization processes for TTF.  These are contained in the HistoricalDaily and HistoricalMonthly data.frame in the InputObject based on index-level data.  Analytics (indicated by *'s) are also contained in the ActualCloseAnalytics data.frame based on actual subaccounts where available and index-level otherwise.  If a subaccount ID is shown below, the analytics in ActualCloseAnalytics should be sourced from that subaccount rather than indices.  Each sector/field combination is represented as nm.fld, where nm represents the abbreviated sector name and fld represends the abbreviated field name (in parens below):

* Cash Sectors: excess of curve returns (.ert), total returns (.trt), benchmark weight (.wt), OAS* (.oas), OASD* (.oasd), DTS* (.dts), OASC* (.oasc), OAD* (.oad), YTW* (.ytw)
    + Gov Related (govRel)
    + TIPS (tips)
    + IG Corp (corp) (7GHU)
    + MBS (mbs): ZV spread also needed** (7CHU)
    + ABS (abs) (7DHU)
    + CMBS (cmbs) (7EHU)
    + RMBS (rmbs) (7FHU)
    + CLO (clo) (7PHU)
    + US HY (usHy) (7HHU)
    + Leveraged Loans (levL): spread to benchmark also needed* (7JHU)
    + EMD (emd)
    + EMC (emc) (7LHU)
    + EML (eml)
    
* Synthetics: excess of curve returns (.ert), total returns (.trt) {for consistency}, benchmark weight (.wt) {=0}, OAS* (.oas), OASD* (.oasd), DTS* (.dts), OASC* (.oasc), OAD* (.oad)
    + IG CDX (igCdx)
    + iTraxx Main (mainCdx)
    + HY CDX (hyCdx)
    + iTraxx XOver (xoCdx)
    + CDX EM (emCdx)
    + SnrFin (SnrFin)
    
* Rates: % yield changes for each key rate (.yChg), convexity* (.cvx)
    + US Key Rate 6 Month (krd6M)
    + US Key Rate 2 Year (krd2)
    + US Key Rate 5 Year (krd5)
    + US Key Rate 10 Year (krd10)
    + US Key Rate 20 Year (krd20)
    + US Key Rate 30 Year (krd30)
    
* Benchmark -- US Aggregate Index (Agg):  excess of curve returns (.ert), total returns (.trt), benchmark weight (.wt) {=1}, OAS* (.oas), OASD* (.oasd), DTS* (.dts), OASC* (.oasc), OAD* (.oad)

**These are saved in OAS analytic 

Note:  Please build in the flexibility to expand strategy to include other risk buckets with time.

The following gives more details about RiskBuckets:

```{r InputsObjectRiskBuckets, echo=TRUE}

InputObject$RiskBuckets

```

```{r OutputObjectData,echo=FALSE,include=TRUE}

#InputObject Aggregation

#EfficientFrontier
OutputObject <- list()
riskLevels <- seq(from=100, to=300, by=25)
nRL <- length(riskLevels)
#create hypothetical exposures
eCashSectors <- matrix(rnorm(mean = 0, sd = 0.2, n=length(cashSectors)*nRL), nrow = length(cashSectors), ncol= nRL)
eSynthetics <- matrix(rnorm(mean = .05, sd = 0.2, n=length(synthetics)*nRL), nrow = length(synthetics), ncol = nRL)
eRateSectors <- matrix(rnorm(mean = 0, sd = 0.05, n=length(ratesSectors)*nRL), nrow = length(ratesSectors), ncol = nRL)
df.netExposure <- data.frame(rbind(eCashSectors, eSynthetics, eRateSectors))
rownames(df.netExposure) <- c(cashSectors, synthetics, ratesSectors)
df.grossExposure <- sweep(df.netExposure, 1, InputObject$BenchmarkWgt$benchmarkWgt, "+")
df.CTEV <- df.netExposure * 2
df.C2VOL <- df.netExposure * 3
df.er <- data.frame(runif(min = 0, max = 300, n = nRL))
colnames(df.er) <- 'ExpRtn'

OutputObject$EfficientFrontier$netExposure <- df.netExposure
OutputObject$EfficientFrontier$grossExposure <- df.grossExposure
OutputObject$EfficientFrontier$CTEV <- df.CTEV
OutputObject$EfficientFrontier$C2VOL <- df.C2VOL
OutputObject$EfficientFrontier$ER <- df.er
OutputObject$EfficientFrontier$riskLevels <- riskLevels


#OptimalPortfolios -- N rows
netWgt <- df.netExposure[,5]
grossWgt <- df.grossExposure[,5] 
oas <- runif(min = 0, max = 10, n=N)
oad <- runif(min = -.2, max = .2, n=N)
dts <- runif(min = 0, max = 2, n=N)
ctER <- runif(min = 0, max = 10, n=N)
isolTEV <- runif(min = 0, max = 60, n=N)
cTEV <- df.CTEV[,5]
cvar <- runif (min = -1, max = -.20, n=N)
var <- runif (min = -.8, max = -.1, n=N)

vol <- runif(min = 0, max = 600, n=N)
c2vol <- grossWgt * vol


oad[1:(length(cashSectors)+length(synthetics))] <- 0
dts[((length(cashSectors)+length(synthetics))+1):N] <- 0

derGE <- rep(0, N)
derGE[(length(cashSectors)+1):(N-length(ratesSectors))] <- runif(min = 0, max = .20, n=length(synthetics))

cpoMgn <- rep(0, N)
cpoMgn[(length(cashSectors)+1):(N-length(ratesSectors))] <- runif(min = 0.01, max = .02, n=length(synthetics))

exCvg <- runif(min = 0, max = 0.25, n=N)
exCvg[(length(cashSectors)+1):(N-length(ratesSectors))] <- 0

tmp <- cbind(netWgt, grossWgt, oas,oad,dts,ctER,isolTEV,cTEV, vol, c2vol, var, cvar, derGE,cpoMgn,exCvg)
tmp <- rbind(tmp, colSums(tmp))
df.strategic <- data.frame(tmp)
colnames(df.strategic)<- c('netWgt','grossWgt','oas','oad','dts','ctER','isolTEV','cTEV', 'vol', 'c2vol', 'cvar', 'var', 'derGE','cpoMgn','ecvg')
rownames(df.strategic) <- c(cashSectors, synthetics, ratesSectors, 'Total')  

df.combined <- df.strategic

df.current <- df.strategic
df.current$netWgt <- c(InputObject$CurrentWgt$currentWgt, sum(InputObject$CurrentWgt$currentWgt))

df.target <- df.strategic
df.target$netWgt <- c(InputObject$TargetWgt$targetWgt, sum(InputObject$TargetWgt$targetWgt))

OutputObject$OptimalPortfolio$Strategic <- df.strategic
OutputObject$OptimalPortfolio$Combined <- df.combined
OutputObject$OptimalPortfolio$Current <- df.current
OutputObject$OptimalPortfolio$Target <- df.target

save(OutputObject, file = "TTF/OutputObject.RData")

```

##### OutputObject Specification for generateOptimalPortfolios

The top level function generateOptimalPortfolios in FIPortfolioOptimization package generates all the requisite data to populate the PM Tool.  The object will contain information for generating two types of exhibits (data stored in two fields below):

* Portfolio Exposures, Analytics and Characteristics for a given set of inputs, constraints at a single risk level
    + These view will be quite granular, as the user will publish these optimal exposures for execution that day
    + OptimalPortfolio field
    
* Efficient frontiers spanning a range of potential risk levels
    + This view does not require the granularity above. Used to understand how optimal exposures would change as risk levels change
    + EfficientFronter field
    
```{r OutputObjectSummary,echo=TRUE}

summary(OutputObject)

```
#### OptimalPortfolio in OutputObject

For a given set of inputs, constraints (including target risk level), the optimizer generates two optimal portfolios: STRUCTURAL (no views) and COMBINED (STRUCTURAL with views).  Along with the optimal exposures for each portfolio, we also generate analytics and portfolio characteristics associated with each one.  These are generated for a single target risk level (included in InputsObject).

In addition to these two optimal portfolios, the user should also be able to view the same set of analytics and portfolio characteristics based on CURRENT and TARGET exposures (as of previous close).  The analytics and portfolio characteristics are calculated for each of these portfolios based on the set of net of benchmark weights specified in InputObject fields CurrentWgt and TargetWgt.  

The OptimalPortfolio contains 4 data.frames:  

```{r OutputObjectDataOptimalPortfolio,echo=TRUE}

summary(OutputObject$OptimalPortfolio)

```

Each data.frame contains the risk bucket names as rows and the analytics/portfolio characteristics (defined below) as columns.  There is also an additional row (total) that provides aggregate analytics/portfolio characteristics at the overall portfolio level:


```{r OutputObjectDataOptimalPortfolioRowColNames,echo=TRUE}

rownames(df.strategic)
colnames(df.strategic)

```


Below are more details about the analytics and portfolio characteristics included:

* Net of benchmark exposure (netWgt)
* Gross of benchmark exposure (grossWgt)
* OAS (oas)
* OAD (oad)
* DTS (dts)
* Contribution to expected return (ctER)
* Isolated TEV (isolTEV)
* Contribution to TEV (cTEV)
* Volatility (vol)
* Contribution to volatility (c2vol)
* CVaR (cvar)
* VaR (var)
* Derivatives gross exposure (.derGE)
* CPO margin (.cpoMgn)
* Excess coverage (.ecvg)


#### EfficientFrontier in OutputObject

The EfficientFrontier field includes data used to plot efficient frontiers.  These efficient frontiers illustrate to the user how net and gross exposures and contribution to risk across risk buckets change as different levels of overall portfolio risk is selected by the user.  

EfficientFrontier contains four data.frames with the net exposures (netExposure), gross exposures (grossExposure), contribution to TEV (CTEV), contribution to volatility (C2VOL), expected portfolio return (ER), and a vector of risk levels (riskLevels):

```{r OutputObjectEfficientFrontier, echo=TRUE}

summary(OutputObject$EfficientFrontier)


```

For each of the first 3 data.frames, the rows are the risk buckets specified above and the columns are the portfolio risk levels:

```{r OutputObjectEfficientFrontierRowsColumns, echo=TRUE}

rownames(OutputObject$EfficientFrontier$netExposure)


```


The data.frame df.er contains the expected portfolio return for each risk level.  

The portfolio risk levels over which the efficient frontiers are estimated will generally remain fixed.  However, the PM Tool (post MVP version) should allow for the user to modify these in the TUNABLES UI.  We will then need to modify the InputObject to feed in the riskLevels vector to this calculation engine.


#### Cov Estimation

The covariance estimation has the following API

CovMatrixInputs <- calculateVolatilityCorrelationMatrix(InputObject)

CovMatrixInputs has the fields below:

* CovMatrixInputs$CorrMatrix -- this is the correlation matrix
* CovMatrixInputs$VolVector -- this is the volatility vector
  
To enable modifying any component VolVector and CorrMatrix, we separate the matrix multiplication to convert these into a covariance matrix: 

CovMatrix <- calculateCovarianceMatrixFromVolCorrs(CovMatrixInputs)

Stepping back to the calculateVolatilityCorrelationMatrix function inputs, many of the fields in InputObject will not have been specified by the user at the time of the function call.  However, the four fields that are needed by the estimation process are:

* RiskBuckets
* CovMatrixOrder
* HistoricalDaily
* HistoricalMonthly

The method will not look for any other fields in the InputObject.  

After running calculateCovarianceMatrixFromVolCorrs, please assign as follows to complete the InputObject:

InputObject$CovMatrix <- CovMatrix
 
#### Storage and Example OutputObject

This OutputObject should be stored with each saved instance of the PM Tool together with the InputObject that was used to generate it. Additionally, please store the original unmodified CovMatrix as well as the CovMatrix that was used to generate the OutputObject based on the InputObject.

An example OutputObject with illustrative (not actual) data that focuses on the appropriate data structures is also provided along with this document

In addition, we have published a package of libraries (PortfolioOptimization) as illustration of the above.  

